A Large-scale Study of Polymorphism in R

### 2019-04-02 paper/main.tex
a bit more prose in results

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Mon Apr  1 14:48:55 2019
\begin{table}[ht]
\label{tab:classpolyL2}
\centering
\begin{tabular}{lrr}
  \hline
Class & Count & Percentage \\
  \hline
  primitive, tbl, tbl\_df & 1796 & 2.78 \\
  array, primitive & 1379 & 2.14 \\
  data.table, primitive & 777 & 1.20 \\
  POSIXct, POSIXt & 716 & 1.11 \\
  formula, primitive & 639 & 0.99 \\
%  xts, zoo & 588 & 0.91 \\
%  glm, lm & 494 & 0.77 \\
%  primitive, ts & 432 & 0.67 \\
%  mts, primitive, ts & 384 & 0.59 \\
%  huxtable, primitive & 318 & 0.49 \\
   \hline
\end{tabular}
\caption{Top 5 class signatures for class-based polymorphism in L2.}
\end{table}

@@

To see how we might use class information, we should see how often a monomorphic class signature co-occurs with a polymorphic type signature.
We count \NUMPOLYTYPEMONOCLASS argument positions which match this criteria.

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Mon Apr  1 16:50:22 2019
\begin{table}[ht]
\centering
\begin{tabular}{llrl}
  \hline
  Type Sig. & Class Sig. & Count & Percentage \\
  \hline
  list$<$any$>$, list$<$character$>$ & roxy\_block & 38 & 2.80 \\
  list$<$any$>$, list$<$list$>$ & Rd & 25 & 1.84 \\
  list$<$any$>$, list$<$real$>$ & TSdata & 25 & 1.84 \\
  list$<$any$>$, list$<$list$>$ & xml\_nodeset & 24 & 1.77 \\
  list$<$any$>$, list$<$list$>$, list$<$real$>$ & dendrogram & 24 & 1.77 \\
%  list$<$any$>$, list$<$real$>$ & phylo & 23 & 1.69 \\
%  list$<$any$>$, list$<$list$>$, list$<$real$>$, scalar/real & dendrogram & 21 & 1.55 \\
%  list$<$any$>$, list$<$language$>$ & uneval & 18 & 1.33 \\
%  list$<$any$>$, list$<$list$>$ & dendrogram & 18 & 1.33 \\
%  list$<$any$>$, list$<$list$>$ & lazy\_dots & 17 & 1.25 \\
   \hline
\end{tabular}
\caption{Top instances of type-polymorphic but class-monomorphic signatures.}
\end{table}

\AT{I don't think this table is telling us much...}

Top attributes, maybe?

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Mon Apr  1 23:06:59 2019
\begin{table}[ht]
\centering
\begin{tabular}{lrl}
  \hline
attr & count & perc \\
  \hline
\{\}, \{names: \C\} & 16589 & 21.11 \\
  \{\}, \{dim: \I\} & 5010 & 6.38 \\
  \{dim:\I,dimnames:list$<$any$>$\}, \{dim:\I\} & 2264 & 2.88 \\
  \{class:\C,names:\C,row.names:\C\}, \{class:\C,names:\C,row.names:\I\} & 2102 & 2.68 \\
  \{\}, \{srcref:\I\} & 2037 & 2.59 \\
  \{dim:\I,dimnames:list$<$\C$>$\}, \{dim:\I\} & 1740 & 2.21 \\
  \{\}, \{class:\C\} & 1673 & 2.13 \\
  \{\}, \{class:\C,levels:\C\} & 1437 & 1.83 \\
  \{\}, \{class:\C,names:\C\} & 1324 & 1.69 \\
  \{\}, \{class:\C,names:\C,row.names:\I\} & 1251 & 1.59 \\
   \hline
\end{tabular}
\end{table}

A lot of the polymorphism that we see is because we haven't propagated our changes through the attribute types.
Will fix this.

>>>>

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Mon Apr  1 14:48:55 2019
\begin{table}[ht]
\label{tab:classpolyL2}
\centering
\begin{tabular}{lrr}
  \hline
Class & Count & Percentage \\
  \hline
  primitive, tbl, tbl\_df & 1796 & 2.78 \\
  array, primitive & 1379 & 2.14 \\
  data.table, primitive & 777 & 1.20 \\
  POSIXct, POSIXt & 716 & 1.11 \\
  formula, primitive & 639 & 0.99 \\
   \hline
\end{tabular}
\caption{Top 5 class signatures for class-based polymorphism in L2.}
\end{table}

@@

To see how we might use class information, we should see how often a monomorphic class signature co-occurs with a polymorphic type signature.
We count \NUMPOLYTYPEMONOCLASS argument positions which match this criteria, which certainly doesn't explain the much higher number of list polymorphism in L1+.

In our analysis, we did not encounter any substantial patterns with class-based polymorphism.
Class is nonetheless important in R, as the dynamic dispatcher uses class information to select the correct method to call.
In short, a small optional annotation indicating the class of an argument may well be enough.

Let us now turn our attention to attributes.
The top contributors for attribute-based polymorphism can be found in Table~\ref{tab:topattrpoly}.

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Mon Apr  1 23:06:59 2019
\begin{table}[ht]
\label{tab:topattrpoly}
\centering
\begin{tabular}{lrl}
  \hline
Attribute Signature & Count & Percentage \\
  \hline
\{\}, \{names: \C\} & 16589 & 21.11 \\
  \{\}, \{dim: \I\} & 5010 & 6.38 \\
  \{dim:\I,dimnames:list$<$any$>$\}, \{dim:\I\} & 2264 & 2.88 \\
  \{class:\C,names:\C,row.names:\C\}, \{class:\C,names:\C,row.names:\I\} & 2102 & 2.68 \\
  \{\}, \{srcref:\I\} & 2037 & 2.59 \\
  \{dim:\I,dimnames:list$<$\C$>$\}, \{dim:\I\} & 1740 & 2.21 \\
  \{\}, \{class:\C\} & 1673 & 2.13 \\
  \{\}, \{class:\C,levels:\C\} & 1437 & 1.83 \\
  \{\}, \{class:\C,names:\C\} & 1324 & 1.69 \\
  \{\}, \{class:\C,names:\C,row.names:\I\} & 1251 & 1.59 \\
   \hline
\end{tabular}
\end{table}

First, the largest source of attribute-based polymorphism is in optional list names.
The names attribute on a vector or list allows indexing with name rather than numeric position.
Another significant source of this polymorphism is matrices, which can be seen in all the ``dim'' and ``dimnames'' attributes.
